### 2023-05-03 tesis_Luis/Abstract/resumen.tex
updated

\begin{abstract}
    \begin{center}
    {\huge Resumen}
\end{center}

\end{abstract}

>>>>

\begin{abstract}
    \begin{center}
    {\LARGE Desarrollo y calibración del sistema de medición de luminosidad CMS Silicon Pixel Detector}
\end{center}

Se describe la calibraci\'on del lumin\'ometro de conteo de píxeles del CMS realizada mediante el programa de escaneo "Van der Meer" para el experimento CMS en la toma de datos en colisiones protón-protón a $\sqrt{s}=$13.6 TeV en el 2022. El período de recopilación de datos fue el 10 y 11 de noviembre para el fill 8381 del LHC, los datos se procesaron  aplicando una selecci\'on de  m\'odulos "buenos" en el  detector, basandose en un estudio de estabilidad de \'estos. El análisis de los escaneos van der Meer con algunas correcciones específicas arroja en la constante de calibraci\'on $\sigma_{vis}$ un valor de:   $4163.3 \pm 3.3 (\text{stat.}) \pm 13 (\text{syst.})\text{ mb}$.

%Se describe la calibraci\'on de la medici\'on de luminosidad para la corrida 3 del gran colisionador de hadrones (LHC),  realizada mediante el programa de escaneo  "Van der Meer"  para el experimento CMS  durante 10 y 11 de Noviembre del 2022 . La toma de datos  proton-proton a $\sqrt{s}=$13.6 TeV fue con  el detector de pixeles de silicio (Silicon Pixel detector) para el  fill 8381, dichos datos se reprocesaron  aplicando una selecci\'on de  m\'odulos "buenos" en el  detector, basandose en un estudio de estabilidad de \'estos. El an\'alisis de 5 escaneos de Van der Meer presentados en este trabajo arroja en la constante de calibraci\'on $\sigma_{vis}$ un valor de:   $4163 \pm 3 (\text{stat.}) \pm 12 (\text{syst.})\text{ mb}$. 
\end{abstract}

### 2023-05-03 tesis_Luis/Chapter1/chapter1.tex
updated

\begin{equation}
  \mathcal{L}_{inst}\propto K\cdot \int\int \int_{-\infty}^{\infty} \rho_{1}(x,y,s,-s_{0}) \rho_{2}(x,y,s,s_{0})dxdydsds_{0}
    \label{lumi_1}
\end{equation}

@@

For a specific case, asuming equal beams, $\sigma_{1x}= \sigma_{2x} ,\sigma_{1y}= \sigma_{2y}$:

@@

This is the well known expression for the luminosity of two Gaussian beams colliding head-on. It shows how the luminosity depends on the number of particles per bunch and the beam sizes \cite{concept_of_luminosity}.

@@

In order to evaluate the total number of events we use integrated luminosity,it is a measurement of the collected data size, and it is an important value to characterize the performance of an accelerator, defined as \cite{concept_of_luminosity}:

@@

Understanding luminosity with the best possible precision is crucial for physics analyses such as searches for new particles, rare processes, or measurements of known particle properties \cite{lumi_motiv}.
The experimental techniques to determine signal rates are mature enough, where the understanding of acceptances, detector biases, reconstruction efficiencies, or background subtraction is at the sub-percent level, so that the final precision of the physics measurement is dominated by the luminosity uncertainty \cite{lumi_paper_def_and_concept}. Ultimately, the greater the precision of the luminosity measurement, the more physicists can understand their observations and delve into hidden corners beyond our current knowledge.

@@

The Large Hadron Collider (LHC) is the world’s largest and most powerful particle accelerator. It first started up on 10 September 2008, and remains the latest addition to CERN’s accelerator complex. The LHC consists of a 27-kilometre ring of superconducting magnets with a number of accelerating structures to boost the energy of the particles along the way. Inside the accelerator, two high-energy particle beams travel at close to the speed of light before they are made to collide. The beams travel in opposite directions in separate beam pipe where the two tubes kept at ultrahigh vacuum. They are guided around the accelerator ring by a strong magnetic field maintained by superconducting electromagnets, these are built from coils of special electric cable that operates in a superconducting state, efficiently conducting electricity without resistance or loss of energy. This requires chilling the magnets to 271.3°C  a temperature colder than outer space \cite{LHC}.\\

@@

The production of proton beams for the LHC involves a complex series of pre-acceleration stages. To begin with, hydrogen is ionized to produce protons, which are then accelerated in bunches up to 50 MeV within the Linear Accelerator 2 (LINAC2). Subsequently, these proton bunches are routed through three circular accelerators - the Booster, the Proton Synchrotron (PS), and the Super Proton Synchrotron (SPS)  in order to gradually increase their energy levels to 1.4 GeV, 26 GeV, and 450 GeV, respectively. Following these pre-acceleration stages, the proton bunches are injected into the LHC ring where they are accelerated to reach energies of up to 7 TeV per bunch while circulating in opposite directions. This entire sequence of steps defines a single LHC fill, which typically involves $10^{14}$ protons that are grouped into bunches to form the proton beam. Figure \ref{lhc_com} depicts the complete CERN accelerator complex.

@@

Since its inception in 2010 the LHC has undergone three runs and two lengthy shutdowns for maintenance and upgrades. 
Run1 began in 2010 with proton-proton collisions at a center-of-mass energy of $\sqrt{s}=\text{7 TeV}$ and concluded in August 2012 at  $\sqrt{s}=\text{8 TeV}$\cite{LHC_status_2013}. Run 2, which took place at $\sqrt{s}=\text{13 TeV}$, was divided into two stages; the first  occurred from 2015 to 2016, and the second part  was carried out   from 2017 to 2018 \cite{pas_17,pas_18}, efforts are currently underway to further improve the precision of this run. Finally  Run3, which began in 2022 at $\sqrt{s}=\text{13.6 TeV}$ and is currently ongoing. This thesis project aims to contribute to the improvement of the precision of the luminosity for Run 3, where it is expected to obtain an approximate luminosity of $390 fb^{-1}$. \\

The integrated luminosity for each of these runs and years, and their corresponding uncertainty values are presented in table \ref{lumi_values} (luminosity values includes the time from the start of stable beams until the beam is dumped by the LHC), also a visualization of these values measured with CMS is shown in Figure \ref{lumi_per_year_int}, which depicts the delivered luminosity versus time for each year.

\begin{table}[h!]
    \begin{center}
    \caption{Integrated luminosity values recorder by each Run}
	\label{lumi_values}
\begin{tabular}{cccc}
\textbf{Run} & \textbf{Energy $\sqrt{s}$} & \begin{tabular}[c]{@{}c@{}}\textbf{Integrated}\\\textbf{Luminosity}\end{tabular} & \textbf{Precision}    \\ 
\toprule
1(2010-2012) & 8 TeV           & $30 fb^{-1}$                                                                               & 2.5\%                 \\
2(2015-2016) & 13 TeV          & $45.9 fb^{-1}$                                                                             & 1.3\%                 \\
2(2017-2018) & 13 TeV          & $118 fb^{-1}$                                                                              & 2.5\%                 \\
3(2022-2024) & 13.6 TeV        & \begin{tabular}[c]{@{}c@{}}$45 fb^{-1}$(current)\\$390 fb^{-1}$(expected)\end{tabular}                  & \begin{tabular}[c]{@{}c@{}}currently\\estimating  \end{tabular}\\
\end{tabular}
    \end{center}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{comment}
\begin{table}[h!]
\begin{center}
    \caption{Integrated luminosity values recorder by each Run}
	\label{lumi_values2}
\begin{tabular}{ccccc}
\multicolumn{1}{l}{\textbf{Run}} & \textbf{Year} & \textbf{Energy $\sqrt{s}$} & \begin{tabular}[c]{@{}c@{}}\textbf{Integrated}\\\textbf{Luminosity}\end{tabular} & \textbf{Precision}    \\ 
\toprule
\multirow{3}{*}{1}               & 2010          & 7 TeV           & $45 pb^{-1}$                                                                                 & 2.5\%                 \\
                                 & 2011          & 7 TeV           & $6.1 fb^{-1}$                                                                                & 2.4\%                 \\
                                 & 2012          & 8 TeV           & $23.3 fb^{-1}$                                                                               & 2.5\%                 \\ 
\hline
\multirow{2}{*}{2}               & 2015          & 13 TeV          & $4.3 fb^{-1}$                                                                              & 1.6\%                 \\
                                 & 2016          & 13 TeV          & $2.6 fb^{-1}$                                                                              & 1.2\%                 \\ 
\hline
\multirow{2}{*}{2}               & 2017          & 13 TeV          & $49.8 fb^{-1}$                                                                             & 2.3\%                 \\
                                 & 2018          & 13 TeV          & $67.9 fb^{-1}$                                                                              & 2.5\%                 \\ 
\hline
\multirow{2}{*}{3}               & 2022          & 13.6 TeV        & $45 fb^{-1}$                                                                                & currently estimating  \\
                                 & 2023          & 13.6 TeV        & currently recording                                                              & currently estimating 
\end{tabular}
 \end{center}
\end{table}
\end{comment}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
  \begin{center}
  \begin{figure}[ht]
    \centering
    \includegraphics[scale=.28]{Chapter1/int_lumi_per_year.png}
    \caption[CMS Luminosity per year]{Delivered luminosity versus time for Run-1 (2010-2012)\cite{pas_17} and Run-2 (2015-2018)\cite{pas_18} and Run-3 that is still ongoing . Cumulative luminosity versus day delivered to CMS during stable beams for pp collisions at nominal center-of-mass energy.  \cite{wikicern}.}
    \label{lumi_per_year_int}
  \end{figure}
    \end{center}
%% tabla con ano , luminosidad y presicion

>>>>

\begin{equation}
  \mathcal{L}_{inst}\propto K\cdot \int\int\int \int_{-\infty}^{\infty} \rho_{1}(x,y,s,-s_{0}) \rho_{2}(x,y,s,s_{0})dxdydsds_{0}
    \label{lumi_1}
\end{equation}

@@

For a specific case, asuming equal beams, $\sigma_{1x}= \sigma_{2x} \text{ and } \sigma_{1y}= \sigma_{2y}$:

@@

This is the well known expression for the luminosity of two Gaussian beams colliding head-on. It shows how the luminosity depends on the number of particles per bunch and the beam sizes \cite{concept_of_luminosity}.\\

@@

In order to evaluate the total number of events we use integrated luminosity, it is a measurement of the collected data size, and it is an important value to characterize the performance of an accelerator, defined as \cite{concept_of_luminosity}:

@@

Understanding luminosity with the best possible precision is crucial for physics analyses such as searches for new particles, rare processes, or measurements of known particle properties.
The experimental techniques to determine signal rates are mature enough, where the understanding of acceptances, detector biases, reconstruction efficiencies, or background subtraction is at the sub-percent level, so that the final precision of the physics measurement is dominated by the luminosity uncertainty . Ultimately, the greater the precision of the luminosity measurement, the more physicists can understand their observations and delve into hidden corners beyond our current knowledge \cite{luminosity_importance}.

@@

The Large Hadron Collider (LHC) is the world’s largest and most powerful particle accelerator. It first started up on 10 September 2008, and remains the latest addition to CERN’s accelerator complex. The LHC consists of a 27-kilometre ring of superconducting magnets with a number of accelerating structures to boost the energy of the particles along the way. Inside the accelerator, two high-energy particle beams travel at close to the speed of light before they are made to collide. The beams travel in opposite directions in separate beam pipe where the two tubes kept at ultrahigh vacuum. They are guided around the accelerator ring by a strong magnetic field maintained by superconducting electromagnets, these are built from coils of special electric cable that operates in a superconducting state, efficiently conducting electricity without resistance or loss of energy. This requires chilling the magnets to -271.3°C  a temperature colder than outer space \cite{LHC}.\\

@@

The production of proton beams for the LHC involves a complex series of pre-acceleration stages. To begin with, hydrogen is ionized to produce protons, which are then accelerated in bunches up to 50 MeV within the Linear Accelerator 2 (LINAC2). Subsequently, these proton bunches are routed through three circular accelerators - the Booster, the Proton Synchrotron (PS), and the Super Proton Synchrotron (SPS)  in order to gradually increase their energy levels to 1.4 GeV, 26 GeV, and 450 GeV, respectively \cite{lhc_complex}. Following these pre-acceleration stages, the proton bunches are injected into the LHC ring where they are accelerated to reach energies of up to 7 TeV per bunch while circulating in opposite directions. This entire sequence of steps defines a single LHC fill, which typically involves $10^{14}$ protons that are grouped into bunches to form the proton beam. Figure \ref{lhc_com} depicts the complete CERN accelerator complex.

@@

Since its inception in 2010 the LHC has undergone three runs and two lengthy shutdowns for maintenance and upgrades. 
Run1 began in 2010 with proton-proton collisions at a center-of-mass energy of $\sqrt{s}=\text{7 TeV}$ and concluded in 2012 at  $\sqrt{s}=\text{8 TeV}$. Run 2, which took place at $\sqrt{s}=\text{13 TeV}$, was divided into two stages; the first  occurred from 2015 to 2016, and the second part  was carried out   from 2017 to 2018, efforts are currently underway to further improve the precision of this run. Finally  Run3, which began in 2022 at $\sqrt{s}=\text{13.6 TeV}$ and is currently ongoing. This thesis project aims to contribute to the improvement of the precision of the luminosity for Run 3, where it is expected to obtain an approximate luminosity of $450 fb^{-1}$ \cite{
hl-lhc}. \\

The integrated luminosity for each of these runs and years, and their corresponding uncertainty values are presented in table \ref{lumi_values} (luminosity values includes the time from the start of stable beams until the beam is dumped by the LHC), also a visualization of these values measured with CMS is shown in Figure \ref{lumi_per_year_int}, which depicts the delivered luminosity versus time for each year.
% ref. TDR Bril para los 390
%separar por lumpas , poner  referencia, poner periodo en vez de Run
%-  en incertidumbre  no conocida, dejar solamente 2022
\begin{table}[h!]
    \begin{center}
    \caption{Integrated luminosity values recorder by each Run}
	\label{lumi_values}
\begin{tabular}{cccc}
\textbf{Period} & \textbf{Energy $\sqrt{s}$} & \begin{tabular}[c]{@{}c@{}}\textbf{Integrated}\\\textbf{Luminosity}\end{tabular} & \textbf{Precision}    \\ 
\toprule
Run 1(2010-2012) \cite{PCC_PAS_12_001} & 8 TeV                   & $30 fb^{-1}$                         & 2.5\%                 \\
Run 2(2015-2016) \cite{lumi_precise_2015_2016} & 13 TeV   & $45.9 fb^{-1}$                     & 1.3\%                 \\
Run 2(2017) \cite{pas_17} & 13 TeV                                                   & $49.8 fb^{-1}$                     & 2.5\%                 \\
Run 2(2018) \cite{pas_18} & 13 TeV                                                  & $67.9 fb^{-1}$                      & 2.5\%                 \\
Run 3(2022) \cite{wikicern} & 13.6 TeV                                            & $42 fb^{-1}$                        &      -                       \\
\end{tabular}
    \end{center}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
  \begin{center}
  \begin{figure}[ht]
    \centering
    \includegraphics[scale=.35]{Chapter1/int_lumi_per_year.png}
    \caption[CMS Luminosity per year]{Delivered luminosity versus time for Run-1 (2010-2012) and Run-2 (2015-2018) and Run-3 that is still ongoing . Cumulative luminosity versus day delivered to CMS during stable beams for pp collisions at nominal center-of-mass energy.  \cite{wikicern}.}
    \label{lumi_per_year_int}
  \end{figure}
    \end{center}
%% tabla con ano , luminosidad y presicion
